Proposal: Loose Grammar Inference for Lossy Compression

### 2012-10-18 proposal.tex
Date
Proofreading fixes

\pagestyle{fancy}
\fancyhead[RO]{Max Wang \& Edward Gan}

@@

\title{Proposal: Loose Grammar Inference for Lossy Compression}
\author{Edward Gan \& Max Wang}
\date{10-17-12}
\maketitle

@@

As described in the introduction, the proposed research shares many of its
goals with the \Sequitur algorithm described in \cite{sequitur}. The authors of
\cite{sequitur} demonstrate how context free grammars offer a means of
simultaneously revealing the structure of data streams and compressing them
effectively. They also develop an effective algorithm based on iteratively
rewriting grammars to keep them small and efficient. \Sequitur yields
intriguing grammar based analyses of texts and musical scores, but an
inspection of its results yields that it isn't able to detect and compress
larger scale structure because of small fluctuations that occur in natural
streams. This was part of our inspiration for trying to achieve both better
compression and better structural analysis by ignoring small scale
fluctuations.

@@

Finally, within the realm of \Sequitur based compression, \cite{nevillphd}
explores a variety of ad-hoc extensions to the \Sequitur algorithm which improve
its compression performance on structured data. These ranged from introducing
domain-specific constraints to their grammar, adding a few steps of
backtracking to their normally greedy grammar formation, to guessing
unifications in attempting to infer recursive grammar rules. Though these may
invalidate some of their theoretical results on the asymptotic performance of
\Sequitur, in practice they seem to have worked and add on nicely into the
grammar inference framework.

The intution that grammar inference can support many detailed policies is made
more formal in \cite{grammarcodes}. The authors classify the properties a
grammar needs to function as a good compressor, and moreover give a set of
reduction rules for putting grammar into the appropriate form. Within this
context, grammar inference algorithms similar to \Sequitur can be described as
simple applications of their reduction rules to different ways of generating
base grammars. Even more examples of the variety of grammar inference schemes
possible under this general framework of reducing grammars can be found in
\cite{efficientgreedy}

@@

Starting with an existing grammar inference algorithm, or perhaps with a simple
variation based on the reduction rules in \cite{grammarcodes}, there are 3
places we might introduce lossiness into the system. We could
\begin{itemize}
  \item perform text-transforms on the input before applying a grammar
    inference algorithms
  \item modify the grammar inference algorithm so that it generates a lossy
    grammar on the fly
  \item take a lossless grammar code and apply lossy transforms to it
\end{itemize}

Of these, the first could provide the biggest potential for taking advantage of
source-specific knowledge of what kinds of data we can ignore. The second
allows the most control over precisely how a grammar can be formed, and would
not be too difficult to build into reduction-rule based frameworks. The third
would be the most general, since it could be applied to any CFG, but may be
difficult to get consistent performance from.

>>>>

\pagestyle{fancy}
\fancyhead[RO]{Edward Gan \& Max Wang}

@@

\title{Proposal: Loose Grammar Inference for Lossy Compression}
\author{Edward Gan \& Max Wang}
\date{October 17, 2012}
\maketitle

@@

As described in the introduction, the proposed research shares many of its
goals with the \Sequitur algorithm described in \cite{sequitur}.  The authors
of \cite{sequitur} demonstrate how context free grammars offer a means of
simultaneously revealing the structure of data streams and compressing them
efficiently.  They also develop an effective algorithm based on iteratively
rewriting grammars to keep them small and efficient. \Sequitur yields
intriguing grammar based analyses of texts and musical scores, inspection of
its results finds that it unable to detect and compress larger scale structure
because of small fluctuations that occur in natural streams.  This was part of
our inspiration for trying to achieve both better compression and better
structural analysis by ignoring small scale fluctuations.

@@

Finally, within the realm of \Sequitur based compression, \cite{nevillphd}
explores a variety of ad-hoc extensions to the \Sequitur algorithm which improve
its compression performance on structured data.  These ranged from introducing
domain-specific constraints to their grammar, adding a few steps of
backtracking to their normally greedy grammar formation, to guessing
unifications in attempting to infer recursive grammar rules.  Though these may
invalidate some of their theoretical results on the asymptotic performance of
\Sequitur, in practice they seem to have worked and add nicely to the grammar
inference framework.

The intution that grammar inference can support many detailed policies is made
more formal in \cite{grammarcodes}.  The authors classify the properties a
grammar needs to function as a good compressor, and moreover give a set of
reduction rules for putting a grammar into the appropriate form.  Within this
context, grammar inference algorithms similar to \Sequitur can be described as
simple applications of their reduction rules to different ways of generating
base grammars.  Even more examples of the variety of grammar inference schemes
possible under this general framework of reducing grammars can be found in
\cite{efficientgreedy}

@@

Starting with an existing grammar inference algorithm, or perhaps with a simple
variation based on the reduction rules in \cite{grammarcodes}, there are three
places we might introduce lossiness into the system.  We could
\begin{itemize}
  \item perform text-transforms on the input before applying a grammar
    inference algorithms;
  \item modify the grammar inference algorithm so that it generates a lossy
    grammar on the fly; or
  \item take a lossless grammar code and apply lossy transforms to it.
\end{itemize}
Of these, the first could provide the biggest potential for taking advantage of
source-specific knowledge of what kinds of data we can ignore.  The second
allows the most control over precisely how a grammar can be formed, and would
not be too difficult, we expect, to build into reduction-rule based frameworks.
The third would be the most general, since it could be applied to any CFG, but
may be difficult to get consistent performance from.
